Composable Generalized Consensus

### 2015-05-14 paper/paper.tex
Work on the presentation of CGC (section 3)

\title{Changing Flights In The Air\\ (Abortable Chain Agreement)}

@@

\begin{comment}
Abortable Chain Agreement builds on several recent advances in SMR
algorithms: ACA allows acceptors to only eventually reach agreement,
like in Generalized Paxos \cite{Lamport05GeneralizeConsensus},
boosting performance when most commands commute; ACA allows fast
execution of read commands like in Generalized Lattice Agreement
\cite{FalerioETAL12GeneralizedLatticeAgreement}; ACA implementations
can tolerate $f$ faults with only $f+1$ replicas and replace crashed replicas
upon changing round, as in Vertical Paxos  \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}.  
\end{comment}

@@

\section{Chain Agreement}
\label{sec:ca}

We consider a set of asynchronous servers communicating by message passing in an
asynchronous network. Servers are not necessarily sequential processes. We
describe a behavior of the system using an infinite sequence of states, where a
state is a function from variable names to values (i.e. a valuation
of variables).  We specify sets of allowed behaviors using an initial predicate,
a next-state relation (using unprimed and primed variables) and a fairness
condition, all in the style of TLA+ \cite{Lamport02SpecifyingSystems}. However,
for succinctness, our specification are rather informal, eluding non-essential
details.

Our specifications all contain by default a state variable $crashed$ and
$Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
executes the algorithm that is assigned to it until its identifier is added to
the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
server is in the set $crashed$, all its actions are disabled.  In an execution,
a server that is never added to the set $crashed$ is said correct.

We consider a computing service exposing a set of commands $C$. The
service has a sequential specification consisting of an initial state
and a deterministic transition relation in which every command
atomically updates the state of the service and produces an output. We
assume that duplicate commands have no effect on the state of the
service and return the same output as their first occurrence.

In a traditional SMR algorithm, the servers can be seen has playing one or
several of three roles: \emph{proposers}, which propose new commands to execute,
\emph{acceptors}, which accept and order commands in a sequence, and
\emph{learners}, which learn about the growing sequence of commands to
execute \cite{lamport2001paxos}. Often, but not necessarily, every replica server plays the
three roles of proposer, acceptor, and learner: clients of the service
send their commands to some replica servers; replica servers submit
the received commands to the SMR protocol in the role of proposers;
replica servers respond to commands by executing the commands learned
in the role of learners; replica servers participate in command
ordering in the role of acceptors.  An SMR algorithm guarantees that
the replicas learn the same sequences of commands, possibly up to the
reordering of commuting commands. Therefore, every learner can respond to client
requests using its last learned sequence of commands, and clients will receive
the same responses as the ones produces by a centralized, non-replicated,
service.

Chain Agreement precisely specifies the allowed behaviors of the roles of \emph{proposers}
and \emph{learners}, abstracting over the behavior of acceptors.

\begin{comment}
We suppose that each server can play a number of \emph{roles}.
Chain Agreement specifies the allowed behaviors the roles of \emph{proposers}
and \emph{learners}. Proposers receive commands from the clients of the service
and propose them for execution by the replicated service. Each learner learns about
a growing sequence of commands. Thanks to the sequential specification of the
service, a sequence of commands determines an output for each the included commands. 
In a traditional SMR algorithm, servers can also play the role of
\emph{acceptors}, whose task is to collectively order commands and let the
learners know about the latest sequence of ordered commands.

The point of Chain Agreement is to implement a replicated service by having every learner learn about a growing sequence of commands, which
represent an execution of the sequential specification of the service.
Crucially, at all times, the sequences of commands learned by the
different learners must form a chain in the prefix order on sequences.
Moreover, for every command $c$ proposed by a correct proposer, every
correct learner must eventually learn a sequence containing $c$.  Once
a learner learns about a sequence containing a command $c$, it can
determine the output produced by $c$, using the sequential
specification of the service, and respond to the client that issued
the command $c$.
\end{comment}

@@

We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
Agreement rounds}, where each round makes progress ordering requests or aborts
its execution and passes the baton to the next round in the sequence.
An algorithm composed of a sequence of ACA rounds is called an \emph{ACA
algorithm}.

On top of the traditional roles of proposer, acceptor, and learner, ACA
introduces the new role of \emph{switcher} and a new component, the
\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
acceptor, learner, and switcher can change in each rounds, allowing for
reconfiguration. 
In normal circumstances, the proposers, acceptors,
and learners of an ACA round work together to implement Chain Agreement.
However, if the environment becomes unfavorable to a running ACA round, the
scheduling policy can instruct the round to abort its execution and pass the
baton, using the switchers, to a new round.

@@

The scheduling policy determines when to abort an ACA round and which new ACA
round should take over. The scheduling policy must inform the switchers about the
identity of the next round and must guarantee that all switchers receive the
same information.  In this paper we abstract over the scheduling policy, assuming
that a round that aborts passes the baton a nondeterministically chosen new
round. A reliable scheduling-policy component can realistically be implemented
as a replicated state machine running on a large number of servers otherwise
dedicated to other tasks
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
 in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
resilience of the service by replacing faulty servers upon round change.

@@

ACA specifies the allowed behaviors of the proposers, learners, and switchers
\emph{of a single round}. Similarly to
Chain Agreement, we use the state variables $crashed$, $proposed$,
and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
we use the state variables $inits$, containing the abort sequences
received from the previous round, and $aborts$, containing the abort
sequences transfered to the next round.
Two consecutive ACA rounds are composed by letting the $aborts$ variable of the
first round be the $inits$ variable of the second round.

@@

\begin{thm}[Composition Theorem]
    \label{thm:comp}
    In every execution of a succession of ACA rounds, the sequence of valuations
    of the variable $proposed$ and, for every learner $l$, of the variables
    $learned\left[ l \right]$ satisfies Chain Agreement.
\end{thm}
\Cref{thm:comp} can be shown by induction on the number of rounds,
using the following lemma.
\begin{lem}
  \label{lem:comp}
    Consider two rounds numbered $i$ and $i+1$. In every execution of a
    succession of ACA rounds, the sequence of valuations of the variables $inits
    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
\end{lem}
\begin{proof}
  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
  The proofs uses a version of the specification of ACA that has a more
  executable form that presented in this section. Moreover,
  instead of sequences of commands, the Isabelle/HOL specification uses the more
  general notion of command history, presented in the next subsection.
  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
  is also formalized in the Isabelle/HOL. The proof uses a refinement
  mapping from the composition of two consecutive ACA instances to a single ACA
  instance.
\end{proof}

\Cref{thm:comp} guarantees that the successive composition of any
number of ACA rounds implements Chain Agreement.
Therefore, with ACA, one can easily build replicated services that
dynamically adapt to the current operating conditions of the system: a
replicated service has several types of rounds at its disposal, each being
tailored to particular conditions. Each time the conditions change,
the current ACA round aborts and passes the baton to a more 
appropriate type of round, as determined by the scheduling policy. Moreover, a new type
of ACA round can always be devised after the system is deployed and added on the fly to
the existing types of rounds.

@@

Most existing SMR algorithms can easily be obtained as the
composition of one or more ACA rounds.

The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
implies that no algorithm can implement SMR to the letter in practice.
Instead, SMR algorithms guarantee safety (i.e. replica execute the
same sequence of commands) but may not be able to make progress (i.e.
order new commands). To make progress as likely as possible, most SMR
algorithms are structured in consecutive rounds (also called ballots,
views, or epochs). In each round, the algorithm tries to make progress
but cannot be guaranteed to succeed. When the algorithm detects that
progress is no more possible, a new round is started.  The new round
is initialized so as to preserve safety, usually by one of the
acceptors. For example, in Paxos, the leader of a newly started ballot
numbered $i$ determines a value that is \emph{safe at $i$} and
proposes it as first value to the other acceptors of the ballot $i$.
In existing SMR algorithms, new rounds are initialized ad-hoc, which
prevents mixing rounds of two different algorithms.  Instead,
ACA is an abstraction of rounds which establishes a common interface and contract for switching from
one round to the next, enabling switching from any ACA round to any other.

To simplify the task of expressing existing algorithms as the composition of one
or more ACA rounds, we propose two specifications, \emph{conservative ACA
rounds} and
\emph{fast ACA rounds}, which can easily be concretized to obtain executable rounds.
We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
Replication, and Ring Paxos can be obtained by concretizing the specification of
conservative ACA rounds, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast ACA rounds.

Fast and conservative ACA rounds are two specifications, at an
intermediate level of abstraction, which satisfy ACA. Both
specifications specify how the state of the acceptors should be
updated but abstract over the particular communication topology used
to implement the state accesses: the specifications use a global state
which is accessible by all servers in all roles. One can refine fast
or conservative rounds by implementing the state accesses and coordination
using the network, obtaining a concrete algorithm.

Conservative and fast ACA rounds are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
``stopped'' and $hist\left[ a \right]$, the local history of the
acceptor $a$, which can be either a history or the special value $None$.
Initially, $inits$, $aborts$, and, for every learner
$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
value $None$.

@@

On top of the additional guard on the $Accept$ transition present only
in conservative rounds, conservative and fast ACA rounds differ in
their definition of learn quorums, abort quorums, and in their
definition of the $SafeAborts\left( R \right)$ operator.
To concisely define read and learn quorums, we define 
\begin{equation}
    Q\left( f \right) = \aset{as \subseteq Acceptors :
        Card\left(as\right) \geq \left\lfloor f\times Card\left(
    Acceptors \right)\right\rfloor + 1}.
\end{equation}

In conservative ACA rounds, abort and learn quorums are defined such
that the intersection between two learn quorums is nonempty and the
intersection between a learn quorum and an abort quorum is nonempty.
The definition leads to two interesting cases: $LearnQ = AbortQ =
Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
\right)$ is defined as the singleton set containing the maximum
element over the local histories of the acceptors in $R$ which are not
$None$. Note that the maximum exists because conservative rounds
guarantee that the histories of the acceptors form a chain.

In fast ACA rounds, abort and learn quorums are defined such that
the intersection between any two learn quorums is nonempty and the
intersection between any two learn quorums and an abort quorum is
nonempty.  The definition leads to three interesting cases: $LearnQ =
AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
The operator $SafeAborts\left( R \right)$ is defined as the set of
maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
: L \in LearnQ}$.

@@

Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
updated each time a new local history is accepted by the learner $l$ by
inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
definition of a history variable).
\begin{thm}
  \label{thm:rounds}
    Both conservative and fast ACA rounds implement ACA under the refinement
    mapping that consists in substituting $\overline{learned\left[ l \right]}$
    for $learned\left[ l \right]$.
\end{thm}

@@

In this section we show how to modify the rounds of Classic Paxos,
Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
ACA rounds. \Cref{thm:comp} then guarantee that all those different types of
rounds can be used in the same algorithm to yield an implementation of Chain
Agreement.

@@

The rounds of the Classic Paxos algorithm can be modified so as to
obtain an implementation of conservative ACA rounds.  We define
Classic Paxos ACA rounds by specializing the specification of
conservative ACA rounds, giving a concrete protocol for disseminating
the proposed commands to the acceptors and enforcing that the learned
histories form a chain.

@@

The specification of Classic Paxos ACA rounds uses the variables of
the specification of conservative ACA rounds, namely the variables
$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
$status\left[ a \right]$, and $hist\left[ a \right]$, and two
additional arrays $accepted\left[ l \right]\left[ a \right]$ and
$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
an acceptor, and $s$ is a switcher.

@@

ACA rounds.

@@

acceptors of the preceeding round, whereas in ACA rounds, the

@@

  Classic Paxos ACA rounds satisfy Abortable Chain Agreement.

@@

  A Classic Paxos ACA round implements the specification of Abortable

@@

In Chain Replication, acceptors are arranged in a chain whose head is
the leader of the round. As in Classic Paxos, abort histories and
proposed commands are sent to the leader. However, instead of the
leader directly sending its local history to all the other acceptors,
the local history of the leader is
forwarded along the chain from one acceptor to the other until it
reaches the last acceptor, which sends the history to the learners. 
By spreading the load more evenly on the all the servers, Chain Replication
can achieve higher throughput that Classic Paxos.
Conservative ACA rounds can be implemented like with Classic Paxos,
replacing the broadcast of the leader by the propagation of the leader's history
along the chain of servers.

Ring Paxos optimizes Chain Replication by having the leader broadcast
commands to the acceptors using IP multicast and only sending hashes
of the commands along the chain. Ring Paxos round can be modified to implement
conservative ACA rounds in a similar way as for Chain Replication.

@@

Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
Paxos, and fast rounds.
The fast rounds can be transformed into
implementations of fast ACA rounds. Instead of sending their
commands to a leader, as in Classic Paxos, the proposers of fast
rounds directly broadcast their commands to all the acceptors. As long
as commands are received in the same order by a learn quorum of
acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
However, if non-commuting commands are received in different orders by
several acceptors, the local history of the acceptors may not form a
chain anymore and the rounds must abort. The fast rounds of Fast
Paxos implement fast ACA rounds.

Fast Paxos ACA rounds are specified using the same set of variables as for
Classic Paxos, initialized to the same values.
The next-state relation is composed of the following types of
transitions, where the definitions of learn quorums, abort quorums,
and of the $SafeAborts\left( R \right)$ operator are the ones of fast
ACA rounds.
\begin{compactitem}
    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
        and a command $c$, not guarded, inserts $c$ in the set
        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
        all the acceptors.
    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
        receive the message, set $hist\left[ a \right]$ to $h$, and to
        set $status\left[ a \right]$ to ``ready''.
    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
        action is to receive the message, to update $hist\left[ a
        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
        the learners.
    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
        the $Learn\left( l, h \right)$ transitions, for a learner $l$
        and a history $h$, the $Stop\left( s \right)$ transitions, for
        a switcher $s$, $Stop\left( a \right)$ transitions, for an
        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
        for a switcher $s$, and the $Abort\left( s, h \right)$
        transitions, for a switcher $s$ and a history $h$, which are
        all the same as in Classic Paxos ACA rounds.
\end{compactitem}

>>>>

\title{Composable Generalized Consensus}

@@

@@

\section{Composable Generalized Consensus}
\label{sec:cgc}

We consider a set of asynchronous servers communicating by message passing in
an asynchronous network. Servers are not necessarily sequential processes.
We describe a behavior of the system using an infinite sequence of states,
where a state is a function from variable names to values (i.e. a valuation of
variables). We specify sets of allowed behaviors using an initial predicate
and a next-state relation expressed using unprimed and primed variables in the
style of TLA+ \cite{Lamport02SpecifyingSystems}. However, for succinctness, our
specification are rather informal, eluding non-essential details.

Our specifications all contain by default a state variable $crashed$ and
 a $Crash\left( x \right)$ transition which is always enabled. A server $x$ faithfully
executes the algorithm that is assigned to it until its identifier is added to
the variable $crashed$ by a $Crash\left( x \right)$ transitions.  When the identifier of a
server is in the set $crashed$, all its actions are disabled.  In an execution,
a server that is never added to the set $crashed$ is said correct.

We consider a computing service exposing a set of commands $C$. The
service has a sequential specification consisting of an initial state
and a deterministic transition relation in which every command
atomically updates the state of the service and produces an output. We
assume that duplicate commands have no effect on the state of the
service and return the same output as did their first occurrence.

The goal of SMR is to provide a reliable implementation of the computing service despite server crashes.
To do so, an SMR algorithm replicates the execution of the service on several servers. 
In a traditional SMR algorithm, the servers can be seen has playing one or
several of three roles: \emph{proposers}, which propose new commands to execute,
\emph{acceptors}, whose role is to order commands, and
\emph{learners}, which learn about the growing sequence of commands to
execute \cite{lamport2001paxos} by querying the acceptors. Often, but not necessarily, every replica server plays the
three roles of proposer, acceptor, and learner: clients of the service
send their commands to some replica servers; replica servers submit
the received commands to the SMR protocol in the role of proposers;
replica servers respond to commands by executing the commands learned
in the role of learners; replica servers participate in command
ordering in the role of acceptors.  A SMR algorithm guarantees that
the replicas learn the same sequences of commands, possibly up to the
reordering of commuting commands. Therefore, every learner can respond to client
requests using its last learned sequence of commands, and clients will receive
the same responses as the ones produces by a centralized, non-replicated,
service.

\subsection{Histories}

To simplify the specification of SMR algorithms optimizing the execution of commuting commands 
we will use the notion of a command history, or history for short, in our specification of Composable Generalized Consensus.

A command history, abbreviated history, is a data structure that
represents a set of sequences which are the same modulo the reordering
of commuting commands and addition or removal of duplicate commands.
We say that two commands $c_1$ and $c_2$ \emph{commute} when, in every
sequential execution of the service in which $c_1$ and $c_2$ are adjacent,
reversing the order of $c_1$ and $c_2$ does not change any of the outputs,
including those given to $c_1$ and $c_2$.

All the sequences represented by a given history contain the same commands
and determine the same outputs for each command. Therefore, to implement a
replicated service, it is sufficient that learners learn histories instead of
sequences. For a thourough discussion of command histories we refer the reader
to \cite{Lamport05GeneralizeConsensus}. Our Isabelle/HOL formalization \cite{}
abstracts over the concrete representation of histories, assuming only their
necessary properties. Command histories were first introduced in trace theory
\cite{Mazurkiewicz84Semantics}.

Command histories can be built from the empty history using the $\star$
operator, which appends a sequence of commands $cs$ to a history $h$, obtaining
the history $h \star cs$. We can then define define an ordering relation among
histories: $h_1 \leq h_2$ when there exists a sequence $cs$ such that $h_2 =
h_1 \star cs$. 
The maximum over a set of
histories $H$ is denoted by $Max\left( H \right)$ and $Max\left( \aset{}
\right)$ is the empty history.
Histories have the property that every finite set $H$ of 
histories has a \emph{greatest lower bound}, noted $GLB\left(H\right)$.

The major advantage that histories have over sequences is that if a sequence
$cs_1$ is obtained by reordering the commuting commands of $cs_2$, then $h \star
cs_1 = h \star cs_2$. Therefore, in implementations of CGC, servers can agree on
a history even though they chose different orders for some commuting commands.

Generalized Consensus precisely specifies the allowed behaviors of the roles of \emph{proposers}
and \emph{learners}, abstracting over the behavior of acceptors.

@@

We now consider implementing Chain Agreement as a sequence of \emph{Abortable Chain
Agreement rounds}, where each round makes progress ordering requests or aborts
its execution and passes the baton to the next round in the sequence.
An algorithm composed of a sequence of CGC instances is called an \emph{ACA
algorithm}.

On top of the traditional roles of proposer, acceptor, and learner, ACA
introduces the new role of \emph{switcher} and a new component, the
\emph{scheduling policy}. In ACA, the assignment of the roles of proposer,
acceptor, learner, and switcher can change in each rounds, allowing for
reconfiguration. 
In normal circumstances, the proposers, acceptors,
and learners of a CGC instance work together to implement Chain Agreement.
However, if the environment becomes unfavorable to a running CGC instance, the
scheduling policy can instruct the round to abort its execution and pass the
baton, using the switchers, to a new round.

@@

The scheduling policy determines when to abort a CGC instance and which new ACA
round should take over. The scheduling policy must inform the switchers about the
identity of the next round and must guarantee that all switchers receive the
same information.  In this paper we abstract over the scheduling policy, assuming
that a round that aborts passes the baton to a nondeterministically chosen new
round. A reliable scheduling-policy component can realistically be implemented
as a replicated state machine running on a large number of servers otherwise
dedicated to other tasks
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}. We will see
 in \cref{sec:rounds} that, as in Vertical Paxos, a reliable scheduling policy component allows
tolerating $f$ faults with $f+1$ replicas, while maintaining the long-term
resilience of the service by replacing faulty servers upon round change.

@@

ACA specifies the allowed behaviors of the proposers, learners, and switchers
\emph{of a single round}. Similarly to
Chain Agreement, we use the state variables $crashed$, $proposed$,
and, for every learner $l$, $learned\left[ l \right]$.  Additionally,
we use the state variables $inits$, containing the abort sequences
received from the previous round, and $aborts$, containing the abort
sequences transfered to the next round.
Two consecutive CGC instances are composed by letting the $aborts$ variable of the
first round be the $inits$ variable of the second round.

@@

\begin{thm}[Composition Theorem]
    \label{thm:comp}
    In every execution of a succession of CGC instances, the sequence of valuations
    of the variable $proposed$ and, for every learner $l$, of the variables
    $learned\left[ l \right]$ satisfies Chain Agreement.
\end{thm}
\Cref{thm:comp} can be shown by induction on the number of rounds,
using the following lemma.
\begin{lem}
  \label{lem:comp}
    Consider two rounds numbered $i$ and $i+1$. In every execution of a
    succession of CGC instances, the sequence of valuations of the variables $inits
    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies ACA.
\end{lem}
\begin{proof}
  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
  The proofs uses a version of the specification of ACA that has a more
  executable form that presented in this section. Moreover,
  instead of sequences of commands, the Isabelle/HOL specification uses the more
  general notion of command history, presented in the next subsection.
  The Isabelle/HOL formalization of ACA uses the theory of I/O automata, which
  is also formalized in the Isabelle/HOL. The proof uses a refinement
  mapping from the composition of two consecutive ACA instances to a single ACA
  instance.
\end{proof}

\Cref{thm:comp} guarantees that the successive composition of any
number of CGC instances implements Chain Agreement.
Therefore, with ACA, one can easily build replicated services that
dynamically adapt to the current operating conditions of the system: a
replicated service has several types of rounds at its disposal, each being
tailored to particular conditions. Each time the conditions change,
the current CGC instance aborts and passes the baton to a more 
appropriate type of round, as determined by the scheduling policy. Moreover, a new type
of CGC instance can always be devised after the system is deployed and added on the fly to
the existing types of rounds.

@@

Most existing SMR algorithms can easily be obtained as the
composition of one or more CGC instances.

The FLP impossibility result \cite{FischerLynchPaterson83ImpossibilityDistributedConsensusOneFaultyProcess}
implies that no algorithm can implement SMR to the letter in practice.
Instead, SMR algorithms guarantee safety (i.e. replica execute the
same sequence of commands) but may not be able to make progress (i.e.
order new commands). To make progress as likely as possible, most SMR
algorithms are structured in consecutive rounds (also called ballots,
views, or epochs). In each round, the algorithm tries to make progress
but cannot be guaranteed to succeed. When the algorithm detects that
progress is no more possible, a new round is started.  The new round
is initialized so as to preserve safety, usually by one of the
acceptors. For example, in Paxos, the leader of a newly started ballot
numbered $i$ determines a value that is \emph{safe at $i$} and
proposes it as first value to the other acceptors of the ballot $i$.
In existing SMR algorithms, new rounds are initialized ad-hoc, which
prevents mixing rounds of two different algorithms.  Instead,
ACA is an abstraction of rounds which establishes a common interface and contract for switching from
one round to the next, enabling switching from any CGC instance to any other.

To simplify the task of expressing existing algorithms as the composition of one
or more CGC instances, we propose two specifications, \emph{conservative ACA
rounds} and
\emph{fast CGC instances}, which can easily be concretized to obtain executable rounds.
We show that ACA versions of Classic Paxos, Multi-Coordinated Paxos, Chain
Replication, and Ring Paxos can be obtained by concretizing the specification of
conservative CGC instances, and that ACA versions of the fast rounds of Fast Paxos can be obtained by concretizing the specification of fast CGC instances.

Fast and conservative CGC instances are two specifications, at an
intermediate level of abstraction, which satisfy ACA. Both
specifications specify how the state of the acceptors should be
updated but abstract over the particular communication topology used
to implement the state accesses: the specifications use a global state
which is accessible by all servers in all roles. One can refine fast
or conservative rounds by implementing the state accesses and coordination
using the network, obtaining a concrete algorithm.

Conservative and fast CGC instances are specified using the state variables $inits$, $proposed$, for every learner $l$, $learned\left[ l \right]$
and $aborts$, as in ACA, and, for every acceptor $a$, two new variables
$status\left[ a \right]$, which can be either ``idle'', ``ready'', or
``stopped'' and $hist\left[ a \right]$, the local history of the
acceptor $a$, which can be either a history or the special value $None$.
Initially, $inits$, $aborts$, and, for every learner
$l$, $learned\left[ l \right]$ are empty sets, and, for every acceptor $a$,
$status\left[ a \right]$  is ``idle'' and $hist\left[ a \right]$ is the special
value $None$.

@@

On top of the additional guard on the $Accept$ transition present only
in conservative rounds, conservative and fast CGC instances differ in
their definition of learn quorums, abort quorums, and in their
definition of the $SafeAborts\left( R \right)$ operator.
To concisely define read and learn quorums, we define 
\begin{equation}
    Q\left( f \right) = \aset{as \subseteq Acceptors :
        Card\left(as\right) \geq \left\lfloor f\times Card\left(
    Acceptors \right)\right\rfloor + 1}.
\end{equation}

In conservative CGC instances, abort and learn quorums are defined such
that the intersection between two learn quorums is nonempty and the
intersection between a learn quorum and an abort quorum is nonempty.
The definition leads to two interesting cases: $LearnQ = AbortQ =
Q\left( 1/2 \right)$; $LearnQ = \aset{Acceptors}$ and $AbortQ =
\aset{\aset{a} : a \in Acceptors}$. The operator $SafeAborts\left( R
\right)$ is defined as the singleton set containing the maximum
element over the local histories of the acceptors in $R$ which are not
$None$. Note that the maximum exists because conservative rounds
guarantee that the histories of the acceptors form a chain.

In fast CGC instances, abort and learn quorums are defined such that
the intersection between any two learn quorums is nonempty and the
intersection between any two learn quorums and an abort quorum is
nonempty.  The definition leads to three interesting cases: $LearnQ =
AbortQ = Q\left( 2/3 \right)$; $LearnQ = Q\left( 1/2 \right)$ and
$AbortQ = Q\left( 3/4 \right)$; $LearnQ = \aset{Acceptors}$ and
$AbortQ = \aset{\aset{a : a \in Acceptors}}$.
The operator $SafeAborts\left( R \right)$ is defined as the set of
maximal elements over the set $\aset{GLB\left( \aset{hist\left[ a
\right] : a \in R \cap L \wedge hist\left[ a \right]\neq None} \right)
: L \in LearnQ}$.

@@

Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
updated each time a new local history is accepted by the learner $l$ by
inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
definition of a history variable).
\begin{thm}
  \label{thm:rounds}
    Both conservative and fast CGC instances implement ACA under the refinement
    mapping that consists in substituting $\overline{learned\left[ l \right]}$
    for $learned\left[ l \right]$.
\end{thm}

@@

In this section we show how to modify the rounds of Classic Paxos,
Multi-Coordinated Paxos, Chain Replication, Ring Paxos, and Fast Paxos to obtain
CGC instances. \Cref{thm:comp} then guarantee that all those different types of
rounds can be used in the same algorithm to yield an implementation of Chain
Agreement.

@@

The rounds of the Classic Paxos algorithm can be modified so as to
obtain an implementation of conservative CGC instances.  We define
Classic Paxos CGC instances by specializing the specification of
conservative CGC instances, giving a concrete protocol for disseminating
the proposed commands to the acceptors and enforcing that the learned
histories form a chain.

@@

The specification of Classic Paxos CGC instances uses the variables of
the specification of conservative CGC instances, namely the variables
$proposed$, $inits$, $aborts$, the arrays $learned\left[ l \right]$,
$status\left[ a \right]$, and $hist\left[ a \right]$, and two
additional arrays $accepted\left[ l \right]\left[ a \right]$ and
$stop\left[ s \right]\left[ a \right]$, where $l$ is a learner, $a$ is
an acceptor, and $s$ is a switcher.

@@

CGC instances.

@@

acceptors of the preceeding round, whereas in CGC instances, the

@@

  Classic Paxos CGC instances satisfy Abortable Chain Agreement.

@@

  A Classic Paxos CGC instance implements the specification of Abortable

@@

In Chain Replication, acceptors are arranged in a chain whose head is
the leader of the round. As in Classic Paxos, abort histories and
proposed commands are sent to the leader. However, instead of the
leader directly sending its local history to all the other acceptors,
the local history of the leader is
forwarded along the chain from one acceptor to the other until it
reaches the last acceptor, which sends the history to the learners. 
By spreading the load more evenly on the all the servers, Chain Replication
can achieve higher throughput that Classic Paxos.
Conservative CGC instances can be implemented like with Classic Paxos,
replacing the broadcast of the leader by the propagation of the leader's history
along the chain of servers.

Ring Paxos optimizes Chain Replication by having the leader broadcast
commands to the acceptors using IP multicast and only sending hashes
of the commands along the chain. Ring Paxos round can be modified to implement
conservative CGC instances in a similar way as for Chain Replication.

@@

Fast Paxos is composed of two types of rounds: classic rounds, like in Classic
Paxos, and fast rounds.
The fast rounds can be transformed into
implementations of fast CGC instances. Instead of sending their
commands to a leader, as in Classic Paxos, the proposers of fast
rounds directly broadcast their commands to all the acceptors. As long
as commands are received in the same order by a learn quorum of
acceptors or are received in different orders but commute, commands can be learned in two message-propagation delays instead of three in Classic Paxos.
However, if non-commuting commands are received in different orders by
several acceptors, the local history of the acceptors may not form a
chain anymore and the rounds must abort. The fast rounds of Fast
Paxos implement fast CGC instances.

Fast Paxos CGC instances are specified using the same set of variables as for
Classic Paxos, initialized to the same values.
The next-state relation is composed of the following types of
transitions, where the definitions of learn quorums, abort quorums,
and of the $SafeAborts\left( R \right)$ operator are the ones of fast
CGC instances.
\begin{compactitem}
    \item A $Propose\left( p,c \right)$ transition, for a proposer $p$
        and a command $c$, not guarded, inserts $c$ in the set
        $proposed$ and broadcasts the message $\aseq{\quo{prop},c}$ to
        all the acceptors.
    \item An $Init\left( s, h \right)$ transition, for a switcher $s$
        and a history $h$, not guarded, inserts $h$ in the set $inits$ and
        broadcasts the message $\aseq{\quo{init},h}$ to all the acceptors.
    \item A $WakeUp\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``idle'' and $a$
        can receive an $\aseq{\quo{init},h}$ message. Its effect is to
        receive the message, set $hist\left[ a \right]$ to $h$, and to
        set $status\left[ a \right]$ to ``ready''.
    \item An $Accept\left( a \right)$ transition, for an acceptor $a$,
        is enabled when $status\left[ a \right]$ is ``ready'' and $a$
        can receive a $\aseq{\quo{prop},c}$ message. The effect of the
        action is to receive the message, to update $hist\left[ a
        \right]$ to $hist\left[ a \right]\star c$, and to broadcast
        the message $\aseq{\quo{ack},hist\left[ a \right]\star c}$ to
        the learners.
    \item The $RcvAck\left( l \right)$ transitions, for a learner $l$,
        the $Learn\left( l, h \right)$ transitions, for a learner $l$
        and a history $h$, the $Stop\left( s \right)$ transitions, for
        a switcher $s$, $Stop\left( a \right)$ transitions, for an
        acceptor $a$, the $RcvStopped\left( s \right)$ transitions,
        for a switcher $s$, and the $Abort\left( s, h \right)$
        transitions, for a switcher $s$ and a history $h$, which are
        all the same as in Classic Paxos CGC instances.
\end{compactitem}

As for Classic Paxos CGC instances, Fast Paxos CGC instances implement ACA
under the refinement mapping consisting projecting the state of the
round onto the variables $proposed$, $inits$, $aborts$, and, for every
learner $l$, $learned\left[ l \right]$.
